Description of radiative transfer submodels

### 2019-09-17 analysis/paper/submodel-rtm.Rmd
Start thinking about crown area
Add two-stream submodel code

Following @clm45_note, forward- ($\nu$) and backscattering of leaves or stems ($\omega$) are defined as a function of reflectance ($R$) and transmittance ($T$) of the corresponding materials:

$$ \nu = R + T $$
$$ \omega = \frac{R + T + \frac{1}{4} (R-T)(1 - \chi)^2}{2 (R+T)} $$

@@

$$
\tau_i = -\exp(-L_i (\phi_1 + \phi_2))
\left(
\phi_1^2 L_i^2 \exp(\phi_1 L_i) * E_i(-\phi_1 L_i) +
\phi_1 L_i - 1
\right)
$$

@@

## Finite canopy radius

@@

The crown area index ($C_i$) affects the following terms in the following ways:

$$ \tau(\psi)_i = (1 - C_i) + C_i \exp(-K(\psi)_i \frac{L_i}{C_i}) $$

$$
\tau_i = (1 - C_i) - 
C_i \exp(-\frac{L_i}{C_i} (\phi_1 + \phi_2))
\left(
\frac{\phi_1^2 L_i^2}{C_i^2} \exp(\phi_1 \frac{L_i}{C_i}) * E_i(-\phi_1 \frac{L_i}{C_i}) +
\phi_1 \frac{L_i}{C_i} - 1
\right)
$$

@@

theta <- 15
S0_beam <- 0.8
S0_diff <- 1 - S0_beam
alb_ground <- 0.1

>>>>

Following @clm45_note, forward- ($\nu$) and backscattering ($\omega$) of canopy elements (leaves or stems) are defined as a function of those elements' reflectance ($R$) and transmittance ($T$):

$$ \nu_x = R_x + T_x $$
$$ \omega_x = \frac{R_x + T_x + \frac{1}{4} (R_x-T_x)(1 - \chi)^2}{2 (R_x+T_x)} $$

@@

$$
\tau_i = -\exp(-L_i (\phi_1 + \phi_2))
\left(
\phi_1^2 L_i^2 \exp(\phi_1 L_i) E_i(-\phi_1 L_i) +
\phi_1 L_i - 1
\right)
$$

@@

$$
Y = 
\begin{bmatrix}
S_0 a_{ground} () \\
S_1 r(\psi)_1 \left[ 1 - r_0 r_1 (1 - \alpha_0) (1 - \tau_0) (1 - \alpha_1) (1 - \tau_1) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) \\
S_1 \left[ 1 - r_1 r_2 (1 - \alpha_1) (1 - \tau_1) (1 - \alpha_2) (1 - \tau_2) \right] (1 - \tau(\psi)_1) (1 - \alpha_1) (1 - r(\psi)_1) \\
... \\
S_i r(\psi)_i \left[ i - r_{i-1} r_i (i - \alpha_{i-1}) (i - \tau_{i-1}) (i - \alpha_i) (i - \tau_i) \right] (i - \tau(\psi)_i) (i - \alpha_i) \\
S_i \left[ i - r_i r_{i+1} (i - \alpha_i) (i - \tau_i) (i - \alpha_{i+1}) (i - \tau_{i+1}) \right] (i - \tau(\psi)_i) (i - \alpha_i) (i - r(\psi)_i) \\
... \\
S_m r(\psi)_m \left[ m - r_{m-1} r_m (m - \alpha_{m-1}) (m - \tau_{m-1}) (m - \alpha_m) (m - \tau_m) \right] (m - \tau(\psi)_m) (m - \alpha_m) \\
S_m \left[ m - r_m r_{m+1} (m - \alpha_m) (m - \tau_m) (m - \alpha_{m+1}) (m - \tau_{m+1}) \right] (m - \tau(\psi)_m) (m - \alpha_m) (m - r(\psi)_m) \\
SW_{sky}
\end{bmatrix}
$$

Diffuse transmittance.

plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Diffuse transmittance fraction")
legend("topright", as.character(orient),
       title = "Orient factor",
       lty = "solid",
       col = colors)
purrr::walk2(
  orient, colors,
  ~curve(ts_tau_diffuse(x, .x, lr = 0.056, lt = 0.045, cai = 1), 0, Lmax, add = TRUE, col = .y)
)

Here it is compared with the multi-scatter model.

orient_sub <- c(-0.5, 0, 0.5)
colors_sub <- colors[c(1, 3, 5)]
plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Diffuse transmittance fraction")
legend("topright", as.character(orient_sub),
       title = "Orient factor",
       lty = "solid",
       col = colors_sub)
legend("top", c("multi-scatter", "two-stream"),
       title = "RTM",
       lty = c("dashed", "solid"))
purrr::walk2(
  orient_sub, colors_sub,
  ~curve(ms_tau_diffuse(x, .x), 0, Lmax, add = TRUE, col = .y, lty = "dashed")
)
purrr::walk2(
  orient_sub, colors_sub,
  ~curve(ts_tau_diffuse(x, .x, lr = 0.056, lt = 0.045, cai = 1), 0, Lmax, add = TRUE, col = .y)
)

# Finite canopy radius

@@

In both models, the transmittance of direct radiation corrected for crown area index ($\tau(\psi)_i^*$) is calculated as follows:

$$ \tau(\psi)_i^* = (1 - C_i) + C_i \exp(-K(\psi)_i \frac{L_i}{C_i}) $$

```{r}
tau_psi_star <- function(L, C, theta, orient) {
  k <- kfunction(theta, orient)
  1 - C + C * exp(-k * L / C)
}
curve(tau_psi_star(1, x, 0, 0), 0, 1,
      ylim = c(0, 1),
      xlab = "Crown area index", y = expression(tau(psi) ^ "*"))
curve(tau_psi_star(2, x, 0, 0), 0, 1, col = 2, add = TRUE)
curve(tau_psi_star(3, x, 0, 0), 0, 1, col = 3, add = TRUE)
curve(tau_psi_star(4, x, 0, 0), 0, 1, col = 4, add = TRUE)
```

The effect of the crown area index on diffuse transmittance varies by model.
In the **two-stream** model, the crown area index affects the average diffuse optical depth ($\bar{\mu}$) as follows:

$$
\bar{\mu}^* = \frac{-L_i}{\log(1 - C_i + C_i \exp( -L_i / \bar{\mu} ))}
$$

(Note that if $C_i = 1$, this simplifies to $\bar{\mu}^* = \bar{\mu}$).

In the **multiple-scatter** model, if we define the diffuse transmittance expression ($\tau_i$) above as a function of leaf area index, $T_i(L)$,
then the diffuse transmittance corrected for crown area ($\tau_i^*$) becomes:

$$
\tau_i^* = (1 - C_i) + C_i T_i\left( \frac{L_i}{C_i} \right)
$$

Comparing the diffuse radiation response of the two models:

cai <- c(0.1, 0.25, 0.5, 1)
colors_sub <- c("red1", "red2", "red3", "black")
plot(c(0, Lmax), c(0, 1), type = "n",
     xlab = "Leaf area index",
     ylab = "Diffuse transmittance fraction")
legend("bottomleft", as.character(cai),
       title = "CAI",
       lty = "solid",
       col = colors_sub,
       inset = c(0, 0.2))
legend("bottomleft", c("MS", "2S"),
       title = "RTM",
       lty = c("dashed", "solid"))
purrr::walk2(
  cai, colors_sub,
  ~curve(ms_tau_diffuse(x, 0, .x), 0, Lmax, add = TRUE, col = .y, lty = "dashed")
)
purrr::walk2(
  cai, colors_sub,
  ~curve(ts_tau_diffuse(x, 0, lr = 0.056, lt = 0.045, cai = .x), 0, Lmax, add = TRUE, col = .y)
)

@@

theta <- 15
S0_beam <- 0.8
S0_diff <- 1 - S0_beam
alb_ground <- 0.1
```
